---
title: "Drought in Texas"
author: "YICHUN LU"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
# Introduction
Texas is no stranger to drought: the state has faced several droughts in recent decades, including the most recent and most severe drought that began in 2011. Texas has faced increasing demand for water to meet the needs of its rapidly growing population. Moreover, climate change has led to Texas experiencing an increase in drought events since 2011, which has intensified the for an impactful solution for state’s water situation. 

Texas is an enormous body of land, and it is difficult to conceptualize the degree of drought without a visual representation. In particular, spatial and temporal visualization can improve our knowledge on the intensity of the drought’s implication on the area throughout Texas. For visualization, the manner of the exploratory focus should be clearly investigated, where factors causing drought range from “Temperature” to “Precipitation” will be analyzed.

Analysis on the relationships between factors of drought, and the spread of the drought area, will be investigated.


# Required packages and data

Please install and load the required packages first.
```{r package loading, echo=FALSE}
# packages installment
library(rmarkdown)
library(gridExtra)
library(ggplot2)
library(rgdal)
library(dplyr)
library(DT)
library(leaflet)
library(weathermetrics)
library(xts)
library(htmlwidgets)
library(widgetframe)
library(dygraphs)
library(zoo)
knitr::opts_chunk$set(cache=TRUE)
```

The boundary shapefile of Texas is one of TIGER products which is downloaded from Census Bureau's MAF/TIGER database.

```{r shpefile loads,echo=FALSE}
# load texas shp file
shp <- readOGR(dsn=path.expand("data/bnd_tx/txdot-counties/txdot-2015-county-generalized_tx.shp"))
knitr::opts_chunk$set(cache=TRUE)
```

In this project, I used drought severity classification data downloaded from United States Drought Monitor (The National Drought Mitigation Center,2017). It is a weekly map that the data is calculated based on measurements of climatic, hydrologic and soil conditions as well as reported impacts and observations from contributors around the country. The time series data ranged from Janaury 5 2010 to November 28 2017.

```{r us drought monitor data loading, message=FALSE, warning=FALSE}
# data1 us monitor data loading
tx_7y <- read.csv('data/tx_7y.csv')
tx_7y_1 <- read.csv('data/tx_7y.csv')
tx_1031 <- read.csv('data/tx_1031.csv')
tx <- as.data.frame(tx_1031)
shp@data = data.frame(shp@data, tx_1031[match(shp@data[["FIPS"]], tx_1031$FIPS),])
# show the data table
datatable(tx_7y_1,style = 'default',options = list(pageLength = 5))
```

Unluckily, the index for each county by state can be downloaded just in single day, therefore, I choosed the data in October 31 as my sample. This is an interactive map which shows the drought spatial distribution in October 31.

```{r interactive map of spatial distribution}
# Transform spatial polygon data frame 
shp_loglat <- spTransform(shp,CRS("+proj=longlat +datum=WGS84"))

D0 <- shp[shp@data$D0!=0,]
D0_loglat <- spTransform(D0, CRS("+proj=longlat +datum=WGS84"))

D1 <- shp[shp@data$D1!=0,]
D1_loglat <- spTransform(D1, CRS("+proj=longlat +datum=WGS84"))

D2 <- shp[shp@data$D2!=0,]
D2_loglat <- spTransform(D2, CRS("+proj=longlat +datum=WGS84"))

# Use leaflet to visualize the drought distribution of each county 
leaflet(shp@bbox) %>%
  addTiles()%>%
    addPolygons(data=shp_loglat,
                weight = 1,
                color = "grey",
                fillColor = "grey",
                stroke = T,
                fillOpacity = 0.1)%>%
  addPolygons(data=D0_loglat,
              label=paste(D0_loglat$CNTY_NM," (D0 =",D0_loglat$D0,")"),
              weight = 2,
              color = "yellow",
              fillColor = "yellow",
              stroke = F,
              fillOpacity = 0.4)%>%
  addPolygons(data=D1_loglat,
              label=paste(D1_loglat$CNTY_NM," (D1 =",D1_loglat$D1,")"),
              weight = 3,
              color = "orange",
              fillColor = "orange",
              stroke = F,
              fillOpacity = 0.8)%>%
  addPolygons(data=D2_loglat,
              label=paste(D2_loglat$CNTY_NM," (D2 =",D2_loglat$D2,")"),
              weight = 4,
              color = "red",
              fillColor = "red",
              stroke = F,
              fillOpacity = 1)%>%
  addMiniMap()%>%
  frameWidget(height = 500)
```

Addtionally, I used the data downloaded from US Climate Division Dataset, including temperature, precipitation and Palmer Drought Severity Index from 2000 to October 2017.

```{r}
# data2 temp$precip data loading
min_mean <- read.csv('data/minmean.csv')
max_mean <- read.csv('data/maxmean.csv')
precip <- read.csv('data/precip.csv')
drought <- read.csv('data/drought.csv')
total <- read.csv('data/total.csv')

# Temperature format transfromation
total$mean.temp <- fahrenheit.to.celsius(total$mean.temp,2)

# inch to cm
total$precip <- total$precip*2.54

# show the data table
datatable(data = total,
          options = list(pageLength = 12),
          colnames = c('Month', 'Mean Temperature (C)', 'Precipitation (cm)', 'Palmer Drought Index'))

```

# Results

## Visualization of drought 

### Drought Percent area in Texas from 2010 to 2017
```{r one-year percent area}
# convert table and date formet 
as.data.frame(tx_7y)
tx_7y$ReleaseDate <- as.Date(as.character(tx_7y$ReleaseDate),format="%Y%m%d");

# chart of one-year drought percent area
ggplot(tx_7y, aes(ReleaseDate, tx_7y$D0)) + geom_area() + geom_area(aes(x = ReleaseDate, y=tx_7y$D1), fill='yellow') + geom_area(aes(x = ReleaseDate, y=tx_7y$D2), fill='orange') + geom_area(aes(x = ReleaseDate, y=tx_7y$D3), fill='red') + xlab("Date") + ylab("Percent Area") 
```

### Sptial distribution of drought in Texas in October 31 2017
```{r visualization the drought area with different drought degree}
# visualization the drought area in D1 degree_ moderate drought
shp2 <- fortify(shp)
shp$id <- row.names(shp)
shp2 <- left_join(shp2,shp@data,by="id")
g0 <- ggplot(data = shp2,
                  aes(x=long,y=lat,group=group,fill=D0))+ geom_polygon() + coord_map() + scale_fill_gradient(low = 'white',high = 'red')

ggplot0<- g0+ theme(legend.position = "right") + theme(axis.ticks = element_blank(), axis.text = element_blank())

# visualization the drought area in D2 degree_ severe drought
g1 <- ggplot(data = shp2,
                  aes(x=long,y=lat,group=group,fill=D1)) + geom_polygon() + coord_map() + scale_fill_gradient(low = 'white',high = 'red')

ggplot1 <- g1+ theme(legend.position = "right")+ theme(axis.ticks = element_blank(), axis.text = element_blank())

# visualization the drought area in D3 Extreme_ severe drought
g2 <- ggplot(data = shp2,
                  aes(x=long,y=lat,group=group,fill=D2))+ geom_polygon() + coord_map() + scale_fill_gradient(low = 'white',high = 'red')

ggplot2 <- g2+ theme(legend.position = "right")+ theme(axis.ticks = element_blank(), axis.text = element_blank())

# build the three graph in one page
grid.arrange(ggplot0,ggplot1,ggplot2,ncol=1)  
```

# Analysis of causes
## Visualization of precipitation and temperature

```{r }
# Convert date column into date format and convert the table into xts format
total[,c("temp_10")] <- total$mean.temp/10
total$X <- as.Date(as.yearmon(total$X))
time <- xts(total[, c(2,3)], order.by=total$X)
# Use dygraph to chart
dygraph(data = time,
        main = "Temperature and precipitation from 2000 to 2017 in Texas",
        height = 100) %>%
  dyRangeSelector(dateWindow = c("2000-01-01", "2017-10-01"))%>%
  dyOptions(stackedGraph = TRUE) %>%
  frameWidget(height =500)
```

## Relationship analysis 
```{r temp & drought}

total%>%
  ggplot(aes(x=X))+
    geom_line(aes(y=Palmer.Drought.Index.Mean),col="grey")+
  geom_line(aes(y=total$precip),col="blue")+
  geom_line(aes(y=mean.temp),col="red")
    
```


# Conclusions
# References

