---
title: "Drought in Texas"
author: "YICHUN LU"
csl: apa.csl
---
# Introduction
Texas is no stranger to drought: the state has faced several droughts in recent decades, including the most recent and most severe drought that began in 2011. Texas has faced increasing demand for water to meet the needs of its rapidly growing population. Moreover, climate change has led to Texas experiencing an increase in drought events since 2011, which has intensified the for an impactful solution for state’s water situation. 

Texas is an enormous body of land, and it is difficult to conceptualize the degree of drought without a visual representation. In particular, spatial and temporal visualization can improve our knowledge on the intensity of the drought’s implication on the area throughout Texas. For visualization, the manner of the exploratory focus should be clearly investigated, where factors causing drought range from “Temperature” to “Precipitation” will be analyzed.

Analysis on the relationships between factors of drought, and the spread of the drought area, will be investigated.


## Required packages and data 

```{r package loading, message=FALSE, warning=FALSE, results='hide'}
# packages installment
library(rmarkdown)
library(gridExtra)
library(ggplot2)
library(rgdal)
library(dplyr)
library(DT)
library(leaflet)
library(weathermetrics)
library(xts)
library(htmlwidgets)
library(widgetframe)
library(dygraphs)
library(zoo)
knitr::opts_chunk$set(cache=TRUE)
```

```{r data loading, message=FALSE, warning=FALSE, results='hide'}
# data1 us monitor data loading
tx_1y <- read.csv('data/tx_1y.csv')
tx_1y_1 <- read.csv('data/tx_1y.csv')
tx_1031 <- read.csv('data/tx_1031.csv')
tx <- as.data.frame(tx_1031)

# data2 temp$precip data loading
min_mean <- read.csv('data/minmean.csv')
max_mean <- read.csv('data/maxmean.csv')
precip <- read.csv('data/precip.csv')
drought <- read.csv('data/drought.csv')
total <- read.csv('data/total.csv')

# load texas shp file
shp <- readOGR(dsn=path.expand("data/bnd_tx/txdot-counties/txdot-2015-county-generalized_tx.shp"))
shp@data = data.frame(shp@data, tx[match(shp@data[["FIPS"]], tx$FIPS),])

knitr::opts_chunk$set(cache=TRUE)
```



# Visualization of drought distribution
## Drought Percent area in Texas
```{r one-year percent area,results='hide'}
# visualization the drought percent area in 1 year
# convert table and date formet 
as.data.frame(tx_1y)
tx_1y$ReleaseDate <- as.Date(as.character(tx_1y$ReleaseDate),format="%Y%m%d");

# datatable visualization
datatable(tx_1y_1,style = 'default',options = list(pageLength = 10))
```

```{r chart_show}
# chart of one-year drought percent area
ggplot(tx_1y, aes(ReleaseDate, tx_1y$D0)) + geom_area() + geom_area(aes(x = ReleaseDate, y=tx_1y$D1), fill='yellow') + geom_area(aes(x = ReleaseDate, y=tx_1y$D2), fill='orange') + geom_area(aes(x = ReleaseDate, y=tx_1y$D3), fill='red') + xlab("Date") + ylab("Percent Area") 
```

```{r leaflet_drought percent}
# Transform spatial polygon data frame 
shp_loglat <- spTransform(shp,CRS("+proj=longlat +datum=WGS84"))

D0 <- shp[shp@data$D0!=0,]
D0_loglat <- spTransform(D0, CRS("+proj=longlat +datum=WGS84"))

D1 <- shp[shp@data$D1!=0,]
D1_loglat <- spTransform(D1, CRS("+proj=longlat +datum=WGS84"))

D2 <- shp[shp@data$D2!=0,]
D2_loglat <- spTransform(D2, CRS("+proj=longlat +datum=WGS84"))

# Use leaflet to visualize the drought distribution of each county 
leaflet(shp@bbox) %>%
  addTiles()%>%
    addPolygons(data=shp_loglat,
                weight = 1,
                color = "grey",
                fillColor = "grey",
                stroke = T,
                fillOpacity = 0.1)%>%
  addPolygons(data=D0_loglat,
              label=paste(D0_loglat$CNTY_NM," (D0 =",D0_loglat$D0,")"),
              weight = 2,
              color = "yellow",
              fillColor = "yellow",
              stroke = F,
              fillOpacity = 0.4)%>%
  addPolygons(data=D1_loglat,
              label=paste(D1_loglat$CNTY_NM," (D1 =",D1_loglat$D1,")"),
              weight = 3,
              color = "orange",
              fillColor = "orange",
              stroke = F,
              fillOpacity = 0.8)%>%
  addPolygons(data=D2_loglat,
              label=paste(D2_loglat$CNTY_NM," (D2 =",D2_loglat$D2,")"),
              weight = 4,
              color = "red",
              fillColor = "red",
              stroke = F,
              fillOpacity = 1)%>%
  addMiniMap()%>%
  frameWidget(height = 500)

```



# Analysis of causes
## Visualization of relation between precipitation and temperature
```{r data surfing}
total$min <- fahrenheit.to.celsius(total$min,2)
total$max <- fahrenheit.to.celsius(total$max,2)
datatable(data = total,
          options = list(pageLength = 12),
          colnames = c('Month', 'Minimum Temperature', 'Maximum Temperature', 'Precipitation', 'Palmer Drought Index'))
```

```{r }
# Convert date column into date format and convert the table into xts format
total$X <- as.Date(as.yearmon(total$X))
time <- xts(total[, c(-1,-5)], order.by=total$X)
# Use dygraph to chart
dygraph(data = time,
        main = "Temperature and precipitation from 2000 to 2017 in Texas",
        height = 100) %>%
  dyRangeSelector(dateWindow = c("2000-01-01", "2017-10-01"))%>%
  dyOptions(stackedGraph = TRUE) %>%
  frameWidget(height =500)
```
## Relationship analysis 
```{r temp & drought}
ggplot(total,aes(y=Palmer.Drought.Index.Mean,x=total$max))+
  geom_point()+
  geom_abline(intercept=0, slope=1)
```

```{r preciptation $ drought}
```


# Results and Conclusions

```{r visualization the drought area with different drought degree,results='hold'}
# visualization the drought area in D1 degree_ moderate drought
shp2 <- fortify(shp)
shp$id <- row.names(shp)
shp2 <- left_join(shp2,shp@data,by="id")
g0 <- ggplot(data = shp2,
                  aes(x=long,y=lat,group=group,fill=D1))+ geom_polygon() + coord_map() + scale_fill_gradient(low = 'white',high = 'red')

ggplot0<- g0+ theme(legend.position = "right") + theme(axis.ticks = element_blank(), axis.text = element_blank())

# visualization the drought area in D2 degree_ severe drought
g1 <- ggplot(data = shp2,
                  aes(x=long,y=lat,group=group,fill=D2)) + geom_polygon() + coord_map() + scale_fill_gradient(low = 'white',high = 'red')

ggplot1 <- g1+ theme(legend.position = "right")+ theme(axis.ticks = element_blank(), axis.text = element_blank())

# visualization the drought area in D3 Extreme_ severe drought
g2 <- ggplot(data = shp2,
                  aes(x=long,y=lat,group=group,fill=D3))+ geom_polygon() + coord_map() + scale_fill_gradient(low = 'white',high = 'red')

ggplot2 <- g2+ theme(legend.position = "right")+ theme(axis.ticks = element_blank(), axis.text = element_blank())

# build the three graph in one page
grid.arrange(ggplot0,ggplot1,ggplot2,nrow=1)  
```

# References

