---
title: "Drought in Texas"
author: "YICHUN LU"
csl: apa.csl
---
# Introduction
Texas is no stranger to drought: the state has faced several droughts in recent decades, including the most recent and most severe drought that began in 2011. Texas has faced increasing demand for water to meet the needs of its rapidly growing population. Moreover, climate change has led to Texas experiencing an increase in drought events since 2011, which has intensified the for an impactful solution for state’s water situation. Texas is an enormous body of land, and it is difficult to conceptualize the degree of drought without a visual representation. In particular, spatial and temporal visualization can improve our knowledge on the intensity of the drought’s implication on the area throughout Texas. For visualization, the manner of the exploratory focus should be clearly investigated, where factors causing drought range from “Temperature” to “Precipitation” will be analyzed.Analysis on the relationships between factors, and the spread of the drought area will illuminate the various causes investigated.


# Required packages and data 
```{r}
# packages installment
install.packages("ggplot2",repos = "http://cran.us.r-project.org")
library(ggplot2)
knitr::opts_chunk$set(cache=TRUE)
# data loading
tx_1y <- read.csv('data/tx_1y.csv')
```



# Visualization of drought
## Drought Percent area in Texas
```{r}
# visualization the drought percent area in 1 year
# convert table and date formet 
as.data.frame(tx_1y)
tx_1y$ReleaseDate<-as.Date(as.character(tx_1y$ReleaseDate),format="%Y%m%d");
ggplot(tx_1y, aes(ReleaseDate, tx_1y$D0))+geom_area()+ geom_area(aes(x = ReleaseDate,y=tx_1y$D1), fill='yellow') +geom_area(aes(x = ReleaseDate,y=tx_1y$D2),fill='orange')+geom_area(aes(x = ReleaseDate,y=tx_1y$D3),fill='red') + xlab("Date")+ylab("Percent Area") 
```

# Analysis of causes
## Visualization of precipitation and temperature
## Relationship analysis 
# Results and Conclusions
# References

